\chapter{List of Publications}
\chaptermark{List of publications}
\label{app:publications}


This dissertation has led to the following communications:

% ----------------------------------------------------------------
\section*{Journal Papers}

\begin{itemize}

\item Christophe Rigaud, Cl{\'e}ment Gu{\'e}rin, Dimosthenis Karatzas, Jean-Christophe Burie and Jean-Marc Ogier. ``Knowledge-driven understanding of images in comic books". In ??? (pending acceptance).
\vspace*{.3cm}

% \item Josep Lladós, Marçal Rusiñol, Alicia Fornés, David Fernández and Anjan Dutta. ``On the Influence of Word Representations for Handwritten Word Spotting in Historical Documents". In International Journal of Pattern Recognition and Artificial Intelligence (IJPRAI), 26(5), August 2012.
\vspace*{.3cm}

\end{itemize}

% ----------------------------------------------------------------
\section*{Book Chapters}

\begin{itemize}

\item Christophe Rigaud, Dimosthenis Karatzas, Jean-Christophe Burie and Jean-Marc Ogier. ``Adaptive contour classification of comics speech balloons". In Graphic Recognition. New Trends and Challenges. Lecture Notes in Computer Science, Vol. 8746, pp. ????, 2014.
\vspace*{.3cm}

\item Christophe Rigaud, Norbert Tsopze, Jean-Christophe Burie and Jean-Marc Ogier. ``Robust frame and text extraction from comic books". In Graphic Recognition. New Trends and Challenges. Lecture Notes in Computer Science, Vol. 7423, pp. 129-138, 2013.
\vspace*{.3cm}

\end{itemize}

% ----------------------------------------------------------------
\section*{Conference Contributions}

\begin{itemize}

\item Cl{\'e}ment Gu{\'e}rin, Christophe Rigaud, Karell Bertet, Jean-Christophe Burie, Arnaud Revel and Jean-Marc Ogier. ``R{\'e}duction de l'espace de recherche pour les personnages de bandes dessin{\'e}es". In the Proceedings of the 19{\`e}me congr{\`e}s national sur la Reconnaissance de Formes et l'Intelligence Artificielle (RFIA), Rouen, France, July, 2014.
\vspace*{.3cm}

\item Christophe Rigaud, Dimosthenis Karatzas, Jean-Christophe Burie and Jean-Marc Ogier. ``Color descriptor for content-based drawing retrieval". In the Proceedings of the 11th IAPR International Workshop on Document Analysis Systems (DAS), pp. 267-271 , Tours, France, April, 2014.
\vspace*{.3cm}

\item Christophe Rigaud, and Cl{\'e}ment Gu{\'e}rin. ``Localisation contextuelle des personnages de bandes dessinées". In the Proceedings of the 13{\`e}me Colloque International Francophone sur l'Ecrit et le Document (CIFED), pp. 367–370, Nancy, France, March 2014.
\vspace*{.3cm}

\item Cl{\'e}ment Gu{\'e}rin, Christophe Rigaud, Antoine Mercier, Farid Ammar-Boudjelal, Karell Bertet, Alain Bouju, Jean-Christophe Burie, Georges Louis, Jean-Marc Ogier and Arnaud Revel. ``eBDtheque: a representative database of comics". In the Proceedings of the 12th International Conference on Document Analysis and Recognition (ICDAR), pp. 1145-1149, Washington DC, USA, August, 2013.
\vspace*{.3cm}

\item Christophe Rigaud, Dimosthenis Karatzas, Joost Van de Weijer, Jean-Christophe Burie and Jean-Marc Ogier. ``An active contour model for speech balloon detection in comics". In the Proceedings of the 12th International Conference on Document Analysis and Recognition (ICDAR), pp. 1240-1244, Washington DC, USA, August, 2013.
\vspace*{.3cm}

\item Christophe Rigaud, Dimosthenis Karatzas, Jean-Christophe Burie and Jean-Marc Ogier. ``Speech balloon contour classification in comics". In the Proceedings of the 10th IAPR International Workshop on Graphics RECognition (GREC), pp. 23-25, Bethlehem, PA, USA, August, 2013.
\vspace*{.3cm}

\item Hoang Nam Ho, Christophe Rigaud, Jean-Christophe Burie and Jean-Marc Ogier. ``Redundant structure detection in attributed adjacency graphs for character detection in comics books". In the Proceedings of the 10th IAPR International Workshop on Graphics RECognition (GREC), pp. 109-113, Bethlehem, PA, USA, August, 2013.
\vspace*{.3cm}

\item Christophe Rigaud, Dimosthenis Karatzas, Joost Van de Weijer, Jean-Christophe Burie and Jean-Marc Ogier. ``Automatic Text Localisation in Scanned Comic Books". In the Proceedings of the 8th International Conference on Computer Vision Theory and Applications (VISAPP), pp. 814-819, Barcelona, Spain, February, 2013.
\vspace*{.3cm}


\item Christophe Rigaud, Norbert Tsopze, Jean-Christophe Burie and Jean-Marc Ogier. ``Extraction robuste des cases et du texte de bandes dessin{\'e}es". In the Proceedings of the 10{\`e}me Colloque International Francophone sur l'Ecrit et le Document (CIFED), pp. 349-360, Bordeaux, France, March 2012.
\vspace*{.3cm}



\end{itemize}

\section*{Websites}

\begin{itemize}
	\item eBDtheque project: \url{http://l3i.univ-larochelle.fr/eBDtheque}
	\item eBDtheque dataset: \url{http://ebdtheque.univ-lr.fr}
	\item Christophe Rigaud's homepage: \url{http://www.christophe-rigaud.com}
\end{itemize}
