\chapter*{Acknowledgement}
\chaptermark{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgement}

First of all I would like to thanks Dr. Simone Marinai and Dr. Apostolos Antonacopoulos to have reviewed this European thesis and for sharing interesting comments and discussions.
I am also grateful to Pr. Dr. Koichi Kise and Pr. Dr. Jean-Philippe Domenger for accepting to be part of the jury and to Dr. Bart Lamiroy to have presided the jury.

Sincere thanks to my supervisors Pr. Dr. Jean-Christophe Burie, Dr. Dimosthenis Karatzas and Pr. Dr. Jean-Marc Ogier for their valuable expertise, friendship, dynamism, review of this thesis and join supervision of this work between the~\CVC~(CVC) of the~\UAB~(UAB), and the~\LIII~(L3i) of the~\ULR~(ULR).
Also, thank you to Remy Mullot for accepting me in his lab (director of the L3i until 2012) and both institutions for providing me all the facilities to carry out this work in a great environment and atmosphere.

I would like to thanks all the people from the Intelligent Media Processing Group of Osaka Prefecture University (Japan) for giving me opportunity of doing research stays with a Manga's team: Pr. Dr. Koichi Kise, Dr. Motoi Iwata, Dr. Weihan Sun, the secretaries and all the master's students including Atsushi Ito and Masaki Tukada which have benefited of a research stay in France.
I have learnt so many things about research, work, life style and food (thank you Hideto!).

Thank you to the CVC members for having welcomed me on several occasions (European doctorate and then joint supervision), especially people from the document analysis group: Josep Llad\'{o}s, Dimosthenis Karatzas, Marçal Rusiñol, Oriol Ramos, Volkmar Frinken, Alicia Fornes, Jon Almazan, Lluis G\'{o}mez, Llu\'{i}s Pere De las Heras, Anjan Dutta, Prass, David Fern\'{a}ndez, Nuria Cirera, Francisco Cruz and Claire Pérez-Mangado from the secretary.
\\

Thank you to the eBDtheque project's team for all the fruitful meetings we had together, Karell Bertet, Jean-Christophe Burie, Arnaud Revel, Alain Bouju, George Louis, Jean-Marc Ogier, Cl\'{e}ment Gu\'{e}rin, Antoine Mercier (annotation tool and dataset's website), Norbert Tsopze and Nam Le Thanh.
Thanks to the L3i lab members that participated to the ground truth day and also to the trainees that contributed to this project.
Sincere thanks to the authors and publishers that have kindly agreed to share part of their works with scientists which allowed us to evaluate, share and make reproducible this three year thesis work.
In alphabetical order: Pascal Boisgibault, Cyb, Fred, Sergio Garcia, Olivier Jolivet, Lamisseb, G\'{e}rald Lubbin, Winsor McCay, Midam, Marion Montaigne, Nicolas Roudier, Alain Saint Ogan, Tr\'ebla and Lewis Trondheim.
Thanks also to their editors: Actes Sud, Ankama, Bac@BD, Clair de Lune, Dargaud, Delcourt, Doc En Stock, Dupuis, Hachette and Studio Cyborga.
Finally, a special thank to the CIBDI\footnote{Cit\'e Internationale de la Bande Dessin\'ee et de l'Image}, Free Public Domain Golden Age Comics and the Department of Computer Science and Intelligent Systems of Osaka Prefecture University who kindly provided material from their personal collection.
\\


Thank you to the people who have taught me how to teach, Vincent Courboulay, Anthony Bourmaud, Karel Bertet, Armelle Prigent, Arnaud Revel and Renaud Peteri.

I would like to thanks the European Doctorate founds of the  University of La Rochelle, European Regional Development Fund, the region Poitou-Charentes (France), the General Council of Charente Maritime (France), the town of La Rochelle (France), the Spanish research projects TIN2011-24631, RYC-2009-05031 and the Sakura project of Campus France for the financial support of this work including the different research stays in Spain and Japan.
Kind thanks to all the volunteer contributors of operating system, software, programming language and libraries related to the completion of this work (Fedora, NetBeans, Sublime Text, Gimp, Inkscape, Firefox, \LaTeX, Python, IPython Notebook, Numpy, SciPy, Matplotlib, OpenCV).
\\

I would like to thanks people that gave me the taste of academic research of image processing during the last year of engineering school and the research master final year project.
You helped me a lot in the application process of this unique and wonderful thesis project, thank you so much.
\\

Thank you to the members of the association of the Ph.D. students of La Rochelle (ADocs) for all the local events we organised together and the help for making and promoting a short film about my research work for young audience.

Grateful thanks to all the Ph.D. candidates, young doctors, engineers and administrative affiliated to the L3i with whom I spend most of the time (Romain, Clément, Guillaume, Cyril, Sébastien, Omar, Van, Giap, Sophea, Dounia, Maroua, Rouaa, Marcela, Bich, Imen, Hind, Phuong, Jane, Muzzamil, Benjamin, Joseph, Bruno, Olivier, Kathy, Caro, Sarah, Dom, Stéphane \& Arnold and those I have forgotten sorry...).

Special thanks to the 121 bis office team (Nat, Mika, To\'{i}, Elo, Nam, Bao, the Roundabout, the Cow and the Space invaders \^{ }\^{ }) for these three years of gentleness, happiness and teasing that we shared together in this unique open space where science and creativity grow on the walls, roof and in the air.
\\

And finally, I would like to infinitely thanks my parents, brother and all my family for their endless love and support at a distance of five hundreds and twenty six kilometres as the crow flies.

Special thanks to my love for having followed me throughout this adventure with patience, dynamism, joy of living and also many contributions to this work (ground truthing, thesis review, etc).





% \begin{itemize}
% 	\item Supervisors + L3i + ULR, CVC (DAG group) + UAB
% 	\item eBDtheque project manager and participant, Antoine Mercier for his programming expertise and the development of ToonShop (annotation tool and demonstrator)
% 	\item eBDtheque dataset material contributors (see acknowledgement section of eBDtheque ICDAR'13 publication)
% 	\item Anh Koi for his first work on the comics at L3i
% 	\item Trainees that worked on the eBDtheque project (sift descriptor, text recognition)
% 	\item Jean-Baptiste Fasquel, Laurent Hardouin et Jean-Louis Boimond, François Chapeau-Blondeau, David Rousseau (engineering school and research master, first works of image processing)
% 	\item Family + love
% 	\item Founding: This work was supported by the European Doctorate founds of the  University of La Rochelle, European Regional Development Fund, the region Poitou-Charentes (France), the General Council of Charente Maritime (France), the town of La Rochelle (France) and the Spanish research projects TIN2011-24631, RYC-2009-05031.
% 	\item All the volunteer contributors of operating system, software, programming language and libraries related to the completion of this work (Fedora, Sublime Text, Gimp, Inkscape, Firefox, \LaTeX, Python, IPython Notebook, Numpy, SciPy, Matplotlib, OpenCV).
% \end{itemize}

\clearpage\thispagestyle{empty}